Extended Data

### 2021-11-15 doc/manuscript/Dow_manuscript_Nature.Rmd
Update Dow_manuscript_Nature.Rmd

###~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
## some stats/ 
species_table <- read.csv(file = "tables_figures/species_table.csv")
n_dendro_trees <- sum(na.omit(species_table[, "n.trees"]))
n_tree_years <- sum(na.omit(species_table[, "n.tree.years"]))
n_cores_focal_sites <- sum(na.omit(species_table[, "n.cores"]))
chronology_table <- read.csv(file = "tables_figures/chronology_table.csv")
n_chronologies <- nrow(chronology_table)  
n_cores_sites <- length(unique(chronology_table[, "Location"]))
n_cores_species <- length(unique(chronology_table[, "Species"]))
n_chronologies_SCBI <- 12
n_chronologies_HF <- 4
```

@@

While the responses of leaf phenology and seasonal _NEE_ to warming spring temperatures have been relatively well-studied[@jeong_phenology_2011; @keenan_net_2014; @buermann_widespread_2018; @ibanez_forecasting_2010; @menzel_european_2006; @menzel_growing_1999], we know very little about how the longest-lived component of fixed carbon in trees, the woody growth, is responding to warming spring temperatures. 
In fact, we know little about the stem-growth phenology of temperate deciduous species, with multi-year records of growth phenology having been published for only one site *(to our knowledge)*[@dorangeville_peak_2021]. <!-- I'm not aware of any others, but we do want to make sure we're not missing anything. klesse_integrating_2016 and martin_hydrometeorology_2017 both do this, but for conifers.-->
The climate sensitivity of woody growth phenology in temperate deciduous trees and its link to annual growth has never been studied *in-situ* (but see Ref.[@zohner_how_2021] for a controlled sapling experiment).

@@

Using dendrometer band measurements taken throughout multiple growing seasons at the Smithsonian Conservation Biology Institute (SCBI; Virginia, USA; n = **125** trees from 2011-2020) and Harvard Forest (Massachusetts, USA; n = **340** trees from 1998-2003), we fit a logistic growth model[@mcmahon_general_2015] to determine the days of year when 25, 50, and 75% annual growth were achieved ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$), peak growing season length ($L_{pgs}= DOY_{75}$-$DOY_{25}$), the maximum daily growth rates ($g_{max}$) and the DOY on which it occurred ($DOY_{g_{max}}$), and total annual increment in diameter at breast height ($\Delta DBH$; Fig. `r schematic_summary`). 
This analysis was performed separately for ring- and diffuse-porous species, which differ in growth phenology[@dorangeville_peak_2021]. These stem-growth milestones were compared to canopy foliage phenology (measured at ecosystem level via remote sensing).

@@

Canopy foliage phenology data for the years 2001-2018 were extracted for SCBI and Harvard Forest from the MCD12Q2 V6 Land Cover Dynamics product (a.k.a. MODIS Global Vegetation Phenology product)[@friedl_mcd12q2_2019]  via Google Earth Engine. Using daily MODIS 2-band Enhanced Vegetation Index data (EVI2) at a spatial resolution of 500m, the product yields the timing of phenometrics (vegetation phenology) over each year, including timing of greenup, midgreenup, and senescence as used in this study.

 <!--Climate -->

>>>>

###~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
## some stats/ 
species_table <- read.csv(file = "tables_figures/species_table.csv")
n_dendro_trees <- sum(na.omit(species_table[, "n.trees"]))
n_dendro_trees_SCBI <- 123 #doing this the lazy way. summed from species_table rather than coded. 
n_dendro_trees_HF <- 340 #doing this the lazy way. summed from species_table rather than coded. 
n_tree_years <- sum(na.omit(species_table[, "n.tree.years"]))
n_cores_focal_sites <- sum(na.omit(species_table[, "n.cores"]))
chronology_table <- read.csv(file = "tables_figures/chronology_table.csv")
n_chronologies <- nrow(chronology_table)  
n_cores_sites <- length(unique(chronology_table[, "Location"]))
n_cores_species <- length(unique(chronology_table[, "Species"]))
n_chronologies_SCBI <- 12
n_chronologies_HF <- 4
```

@@

While the responses of leaf phenology and seasonal _NEE_ to warming spring temperatures have been relatively well-studied[@jeong_phenology_2011; @keenan_net_2014; @buermann_widespread_2018; @ibanez_forecasting_2010; @menzel_european_2006; @menzel_growing_1999], we know very little about how the longest-lived component of fixed carbon in trees, the woody growth, is responding to warming spring temperatures. 
In fact, we know little about the stem-growth phenology of temperate deciduous species, with multi-year records of growth phenology having been published for only one site to our knowledge[@dorangeville_peak_2021]. <!-- I'm not aware of any others, but we do want to make sure we're not missing anything. klesse_integrating_2016 and martin_hydrometeorology_2017 both do this, but for conifers.-->
The climate sensitivity of woody growth phenology in temperate deciduous trees and its link to annual growth has never been studied *in-situ* (but see Ref.[@zohner_how_2021] for a controlled sapling experiment).

@@

Using dendrometer band measurements taken throughout multiple growing seasons at the Smithsonian Conservation Biology Institute (SCBI; Virginia, USA; n = `r n_dendro_trees_SCBI` trees from 2011-2020) and Harvard Forest (Massachusetts, USA; n = `r n_dendro_trees_HF` trees from 1998-2003), we fit a logistic growth model[@mcmahon_general_2015] to determine the days of year when 25, 50, and 75% annual growth were achieved ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$), peak growing season length ($L_{pgs}= DOY_{75}$-$DOY_{25}$), the maximum daily growth rates ($g_{max}$) and the DOY on which it occurred ($DOY_{g_{max}}$), and total annual increment in diameter at breast height ($\Delta DBH$; Fig. `r schematic_summary`). 
This analysis was performed separately for ring- and diffuse-porous species, which differ in growth phenology[@dorangeville_peak_2021]. These stem-growth milestones were compared to canopy foliage phenology (measured at ecosystem level via remote sensing).

@@

Canopy foliage phenology data for the years 2001-2018 were extracted for SCBI and Harvard Forest from the MCD12Q2 V6 Land Cover Dynamics product (a.k.a. MODIS Global Vegetation Phenology product)[@friedl_mcd12q2_2019] via Google Earth Engine. Data were extracted for the pixel containing the NEON tower at each site. Using daily MODIS 2-band Enhanced Vegetation Index data (EVI2) at a spatial resolution of 500m, the product yields the timing of phenometrics (vegetation phenology) over each year, including timing of greenup, midgreenup, and senescence as used in this study. 
 <!--Climate -->
